Appendix to 'Detecting social information in a dense database of infants’ natural

### 2021-05-26 paper/preprint/paper_revised_submission_april2021_apa.Rmd
Intro/Discussion edits, quadratic results, added references

---
title             : "Detecting social information in a dense database of infants’ natural visual experience"
shorttitle        : "Detecting social information"

@@

abstract: |
           The faces and hands of caregivers and other social partners offer a rich source of social and causal information that may be critical for infants' cognitive and linguistic development. Previous work using manual annotation strategies and cross-sectional data has found systematic changes in the proportion of faces and hands in the egocentric perspective of young infants. Here, we examine the prevalence of faces and hands in a longitudinal collection of more than 1700 headcam videos collected from three children along a span of 6 to 32 months of age—the SAYCam dataset [@SAYcam]. To analyze these naturalistic infant egocentric videos, we first validated the use of a modern convolutional neural network of pose detection (OpenPose) for the detection of faces and hands and then applied this model to the entire dataset. First, we found a higher proportion of hands in view than previously reported and a moderate decrease in the proportion of faces in children's view across age. Second, we  found variability in the proportion of faces/hands viewed by different children in different locations (e.g., living room vs. kitchen), suggesting that individual activity contexts may shape the social information that infants experience. Third, we found evidence that children may see closer, larger views of people, hands, and faces earlier in development. These analyses provide new insight into the changes in the social information in view across early development and call for further work that examines their generalizability across populations and their relationship to learning outcomes.

@@

Yet until recently most hypotheses about infants' visual experience have gone untested. 
Though parents and scientists alike have strong intuitions about what infants see, even the viewpoint of a walking child is hard to intuit [@clerkin2017; @franchak2011]. 
By equipping infants and toddlers with head-mounted cameras, researchers have begun to document the infant's egocentric perspective on the world. 
Using these methods, a growing body of work now demonstrates that the viewpoints of very young infants (less than 4 months of age) are indeed dominated by frequent, persistent views of the faces of their caregivers [@Jayaraman2015; @Sugden2014; @Jayaraman2018].

Beyond these early months, infants' motor and cognitive abilities mature, leading to vastly different perspectives on the world. For example, crawlers see fewer faces and hands than do walking children [@sanchez2018postural; @franchak2017see; @kretch2014] as well as different views of objects [@smith2011not]. Further, as infants learn to use their own hands to act on the world, they seem to focus on manual actions taken by their social partners, and their perspective starts to capture views of hands manipulating objects [@Fausey2016]. In turn, caregivers may also start to use their hands with more communicative intent, directing infants' attention by pointing and gesturing to different events and objects during play [@yu2013joint].

Here, we examine the social information present in the infant visual perspective---the presence of faces and hands---by analyzing a longitudinal collection of more than 1700 headcam videos collected from three children along a span of 6 to 32 months of age—the SAYCam dataset [@SAYcam]. In addition to its size and longitudinal nature, this dataset is more naturalistic than those previously used in two key ways. First, recordings were taken under a large variety of activity contexts [@roy2015predicting; @bruner1985role] encompassing infants' viewpoints during both activities outside and inside the home. Even in other naturalistic datasets, the incredible variety in a typical infant's experience has been largely underrepresented (see examples in Figure 1; e.g., riding in the car, gardening, watching chickens during a walk, browsing magazines, nursing, brushing teeth). Second, the head-mounted cameras used in the SAYCam dataset captured a larger field of view than those typically used, allowing a more complete picture of the infant perspective. While head-mounted cameras with a more restricted field of view do represent where infants are foveating most of the time [@yoshida2008; @smith2015contributions], they may fail to capture short saccades to either faces or hands in the periphery, as the timescale of head movements is much longer.

@@

To evaluate OpenPose's performance, we compared its detections to the manually-annotated gold set of frames, calculating precision (hits / (hits + false alarms)), recall (hits / (hits + misses)), and F-score (the harmonic mean of precision and recall).  In our data, for faces, the F-score was `r round(face_performance[3],2)`, with a precision of `r round(face_performance[1],2)` and recall of `r round(face_performance[2],2)`. For hands, the F-score was `r round(hand_performance[3],2)`, with a precision of `r round(hand_performance[1],2)` and recall of `r round(hand_performance[2],2)`. While face and hand detections showed moderately good precision, face detections were overall slightly more accurate than hand detections. In general, hand detections suffered from fairly low recall, indicating that OpenPose likely underestimated the proportion of hands in the dataset.   We also found that restricing our detections to high-confidence face/hand detections (>.5 confidence, default threshold for visualization in OpenPose) was not beneficial -- improving precision but dramatically impairing recall and thus overall performance: the F-score for high-confidence face detections was `r round(face_performance_hc[3],2)`, with a precision of `r round(face_performance_hc[1],2)` and recall of `r round(face_performance_hc[2],2)`; for high-confidence hand detections, the F-score was `r round(hand_performance_hc[3],2)`, with a precision of `r round(hand_performance_hc[1],2)` and recall of `r round(hand_performance_hc[2],2)`)

We suspected that this was in part because children’s own hands were often in view of the camera and unconnected to a pose--a notoriously challenging detection problem [@bambach2015lending]. To assess this possibility, we obtained human annotations for the entire subsample of 9051 frames in which a hand was detected; participants (recruited via AMT) were asked to draw bounding boxes around children’s and adult’s hands. Overall, we found that `r round(mean(hands_missed$child_hand_seg),2)*100`% of missed hand detections were of child hands. When frames with children's hands were removed from the gold set, recall did improve somewhat to `r round(hand_performance_adults[3],2)`. We also observed that children’s hands tended to appear in the lower half of the frames; heatmaps of the bounding boxes obtained from these annotations can be seen in Appendix Figure \ref{fig:handLocation}.

@@

We analyzed the social information in view across the entire dataset, looking specifically at the proportions of faces and hands detected for each child.^[All analyses and preprocessed data files for this paper are available at https://tinyurl.com/detecting-social-info] Data from videos were binned according to the age of the child (in weeks). First, we saw that the proportion of faces in view showed a moderate decrease across this age range (see Figure \ref{fig:faceHandDetailed}), in keeping with prior findings [@Fausey2016]; in contrast, we did not observe an increase in the proportion of hands in view. These effects were quantified with two separate linear mixed-models (see Tables 1 & 2).^[Face/hand detections were binned across each week of filming. Participant's age was converted into months and centered for these analyses. Random slopes for the effect of age by child led to a singular fit and were removed from both analyses; see full model specification in accompanying codebase.]

However, the most striking result from these analyses is a much overall greater proportion of hands in view than have previously been reported [@Fausey2016]. We found this to be true across all ages, in all three children, and regardless of whether we analyzed human annotations (on the 24K random subset, see dotted lines in Appendix Figure \ref{fig:goldSetSanity}) or OpenPose annotations on the entire dataset (see Figure \ref{fig:faceHandDetailed}A). This is notable especially given that OpenPose showed relatively low recall for hands, indicating that this may be an underestimate of the proportion of hands in view. In fact, analysis the human annotations underscores revealed a much higher proportion of hands relative to faces than the automated annotations.

@@

# General Discussion

Here, we analyzed the social information in view in a dense, longitudinal dataset, applying a modern computer-vision model to quantify the hands and faces seen from each of three children's egocentric perspective from 6 to 32 months of age. 
First, we found a moderate decrease across age in the proportion of faces in view in the videos, in keeping with previous work [@Fausey2016]. 
This finding is particularly notable given that, in previous cross-sectional data, this effect seems to be most strongly driven by infants younger than 4 months of age [e.g., @Fausey2016; @Jayaraman2015; @Sugden2014] who see both more frequent and more persistent faces [@Jayaraman2018].  
We also found this to be true when restricting our analyses to full-field faces, suggesting this effect is not driven by a concurrent shift from more full-view to partial-views of faces.

We also found an unexpectedly high proportion of hands in the view of infants, even when restricting the field-of-view to the center field of view the videos to make the viewpoints comparable to those of headcams used in previous work [@Fausey2016]. 
Why might this be the case? One idea is that these videos contain the viewpoints of children not only during structured interactions (e.g., play sessions at home or in the lab) but during everyday activities when children may be playing by themselves or simply observing the actions of caregivers and other people in their environment. During these less structured times, caregivers may move about in the vicinity of the child but not interact with them as directly---leading to views where a person and their hands are visible from a distance, but this person's face may be turned away from the infant or occluded (see examples in Figure 1). Indeed, using the same pose detector on videos from in-lab play sessions, @sanchez2018postural found the opposite trend: slightly fewer hand detections than face detections from 8-16 months of age. Work that directly examines the variability in the social information in view across more vs. less structured activity contexts could further test this idea.

@@

We also used these detailed pose annotations to explore finer-grained changes in how children experience the faces and hands of their caregivers over development. We found that the faces, hands, and people in the infant view tended to become smaller and that faces tended to be farther away and in more variable horizontal positions.  Overall, these data support the idea that the social information in view changes across development as infants become increasingly mobile and independent. As children can navigate the world on their own, they may experience fewer close-up interactions with the caregivers and more bouts of play where they are exploring the objects and things in the environment around them.

More broadly, however, these analyses underscore the importance of how, when, from whom, and what data we sample; these choices become central when we attempt to draw conclusions about the regularities of experience. Indeed, while unprecedented in size, this dataset still has many limitations. These videos only represent a small portion of the everyday experience of these three children, all of whom come from relatively privileged households in western societies and thus are not representative in many ways of the global population. Any idiosyncrasies in how and when these particular families chose to film these videos also undoubtedly influences the variability seen here. And without eye-tracking data, we do not know if children are attending to the social information in their visual field.

Nonetheless, we believe that these advances in datasets and methodologies represent a step in the right direction. The present paper demonstrates the feasibility of using a modern computer vision model to annotate the entirety of a very large dataset (here, >40M million frames) for the presence and size of people, hands, and faces, representing orders of magnitude more data relative to prior work. We propose that the large-scale analysis of dense datasets, collected with different fields of view, cameras and from many different laboratories, will lead to generalizable conclusions about the regularities of infant experience that scaffold learning.

>>>>

---
title             : "A longitudinal analysis of the social information in infants' naturalistic visual experience using automated detections"
shorttitle        : "Longitudinal analysis of social information"

@@

abstract: |
           The faces and hands of caregivers and other social partners offer a rich source of social and causal information that may be critical for infants' cognitive and linguistic development. Previous work using manual annotation strategies and cross-sectional data has found systematic changes in the proportion of faces and hands in the egocentric perspective of young infants. Here, we examine the prevalence of faces and hands in a longitudinal collection of more than 1700 headcam videos collected from three children along a span of 6 to 32 months of age—the SAYCam dataset [@SAYcam]. To analyze these naturalistic infant egocentric videos, we validated the use of a modern convolutional neural network of pose detection (OpenPose) for the detection of faces and hands and then applied this model to the entire dataset. First, we found a higher proportion of hands in view than previously reported and a moderate decrease in the proportion of faces in children's view across age. Second, we found variability in the proportion of faces/hands viewed by different children in different locations (e.g., living room vs. kitchen), suggesting that individual activity contexts may shape the social information that infants experience. Third, we found evidence that children may see closer, larger views of people, hands, and faces earlier in development. These analyses provide new insight into the changes in the social information in view across the first few years of life and call for further work that examines their generalizability across populations and their relationship to learning outcomes.

@@

Yet until recently most hypotheses about infants' visual experience have gone untested. 
Though parents and scientists alike have strong intuitions about what infants see, even the viewpoint of a walking child is hard to intuit [@clerkin2017; @franchak2011]. 
By equipping infants and toddlers with head-mounted cameras, researchers have begun to document the infant's egocentric perspective on the world [@smith2015contributions; @smith2018developing; @franchak2011] and the consequences of this changing view for early learning.
Using these methods, a growing body of work now demonstrates that the viewpoints of very young infants (less than 4 months of age) are indeed dominated by frequent, persistent views of the faces of their caregivers [@Jayaraman2015; @jayaraman2013visual; @jayaraman2017faces; @Sugden2014; @Jayaraman2018].

Beyond these early months, infants' motor and cognitive abilities mature, leading to vastly different perspectives on the world [@iverson2010]. For example, children see fewer faces and hands when crawling than walking or sitting [@luo2020head; @sanchez2018postural; @franchak2017see; @kretch2014; @franchak2019changing; @yamamoto2020transition] as well as different views of objects [@smith2011not; @luo2020head]. Further, as infants learn to use their own hands to act on the world, they seem to focus on manual actions taken by their social partners, and their perspective starts to capture views of hands manipulating objects [@Fausey2016]. In turn, caregivers may also start to use their hands with more communicative intent, directing infants' attention by pointing and gesturing to different events and objects during play [@yu2013joint].

Here, we examine the social information present in the infant visual perspective---the presence of faces and hands---by analyzing a longitudinal collection of more than 1700 headcam videos collected from three children along a span of 6 to 32 months of age—the SAYCam dataset [@SAYcam]. In addition to its size and longitudinal nature, this dataset is more naturalistic than those previously used in two key ways. First, recordings were taken under a large variety of activity contexts [@roy2015predicting; @bruner1985role] encompassing infants' viewpoints during both activities outside and inside the home. Even in other naturalistic datasets, the incredible variety in a typical infant's experience has been largely underrepresented (see examples in Figure 1; e.g., riding in the car, gardening, watching chickens during a walk, browsing magazines, nursing, brushing teeth). Second, the head-mounted cameras used in the SAYCam dataset captured a larger field of view than those typically used, allowing a more complete picture of the infant perspective. While head-mounted cameras with a more restricted field of view do represent where infants are foveating most of the time [@yoshida2008; @smith2015contributions], they may fail when faces or hands appear in children's peripheral vision but are still part of a joint interaction.

@@

To evaluate OpenPose's performance, we compared its detections to the manually-annotated gold set of frames, calculating precision (hits / (hits + false alarms)), recall (hits / (hits + misses)), and F-score (the harmonic mean of precision and recall).  In our data, for faces, the F-score was `r round(face_performance[3],2)`, with a precision of `r round(face_performance[1],2)` and recall of `r round(face_performance[2],2)`. For hands, the F-score was `r round(hand_performance[3],2)`, with a precision of `r round(hand_performance[1],2)` and recall of `r round(hand_performance[2],2)`. While face and hand detections showed moderately good precision, face detections were overall slightly more accurate than hand detections. In general, hand detections suffered from fairly low recall, indicating that OpenPose likely underestimated the proportion of hands in the dataset.  We also found that restricting our detections to high-confidence face/hand detections (>.5 confidence, default threshold for visualization in OpenPose) was not beneficial -- improving precision but dramatically impairing recall and thus overall performance: the F-score for high-confidence face detections was `r round(face_performance_hc[3],2)`, with a precision of `r round(face_performance_hc[1],2)` and recall of `r round(face_performance_hc[2],2)`; for high-confidence hand detections, the F-score was `r round(hand_performance_hc[3],2)`, with a precision of `r round(hand_performance_hc[1],2)` and recall of `r round(hand_performance_hc[2],2)`)

We suspected that this was in part because children’s own hands were often in view of the camera and unconnected to a pose--a notoriously challenging detection problem [@bambach2015lending]. To assess this possibility, we obtained human annotations for the entire subsample of 9051 frames in which a hand was detected; participants (recruited via Amazon Mechanical Turk) were asked to draw bounding boxes around children’s and adult’s hands. Overall, we found that `r round(mean(hands_missed$child_hand_seg),2)*100`% of missed hand detections were of child hands. When frames with children's hands were removed from the gold set, recall did improve somewhat to `r round(hand_performance_adults[3],2)`. We also observed that children’s hands tended to appear in the lower half of the frames; heatmaps of the bounding boxes obtained from these annotations can be seen in Appendix Figure \ref{fig:handLocation}.

@@

We analyzed the social information in view across the entire dataset, looking specifically at the proportions of faces and hands detected for each child.^[All analyses and preprocessed data files for this paper are available at https://tinyurl.com/detecting-social-info] Data from videos were binned according to the age of the child (in weeks). First, we saw that the proportion of faces in view showed a moderate decrease across this age range (see Figure \ref{fig:faceHandDetailed}), in keeping with prior findings [@Fausey2016]; in contrast, we did not observe an increase in the proportion of hands in view. 
These effects were quantified with two separate linear mixed-effect models (see Tables 1 & 2)^[Face/hand detections were binned across each week of filming. Participant's age was converted into months and centered for these analyses. Random slopes for the effect of age by child led to a singular fit and were removed from both analyses; see full model specification in accompanying codebase.] After visualizing the data (see Figure \ref{fig:faceHandDetailed}A), we examined whether linear vs. quadratic terms relating children's age to the proportion of faces/hands detected would provide better fits to the data, and found that this was true in both cases (see Tables 1 & 2), though linear terms also provided a significant fit for faces. Thus, these exploratory results point towards the idea that some children may experience overall more social information in view in the second year of life.

However, the most striking result from these analyses is a much overall greater proportion of hands in view than have previously been reported [@Fausey2016]. We found this to be true across all ages, in all three children, and regardless of whether we analyzed human annotations (on the 24K random subset, see dotted lines in Appendix Figure \ref{fig:goldSetSanity}) or OpenPose annotations on the entire dataset (see Figure \ref{fig:faceHandDetailed}A). This is notable especially given that OpenPose showed relatively low recall for hands, indicating that this may be an underestimate of the proportion of hands in view. In fact, analysis of the human annotations underscores revealed a much higher proportion of hands relative to faces than the automated annotations.

@@

# General Discussion
Here, we analyzed the social information in view in a dense, longitudinal dataset, applying a modern computer-vision model to quantify the hands and faces seen from each of three children's egocentric perspective from 6 to 32 months of age. 
First, we found a moderate decrease across age in the proportion of faces in view in the videos, in keeping with previous work [@Fausey2016; @Jayaraman2015]. 
This finding is particularly notable given that, in previous cross-sectional data, this effect seems to be most strongly driven by infants younger than 4 months of age [e.g., @Fausey2016; @Jayaraman2015; @Sugden2014] who see both more frequent and more persistent faces [@Jayaraman2018].  
We also found this to be true when restricting our analyses to full-field faces, suggesting this effect is not driven by a concurrent shift from more full-view to partial-views of faces.

We also found an unexpectedly high proportion of hands in the view of infants, even when restricting the field-of-view to the center field-of-view to make the viewpoints comparable to those of headcams used in prior work. Why might this be the case? One idea is that these videos contain the viewpoints of children not only during structured interactions (e.g., play sessions at home or in the lab) but during everyday activities when children may be playing by themselves or simply observing the actions of caregivers and other people in their environment. During these less structured times, caregivers may move about in the vicinity of the child but not interact with them as directly---leading to views where a person and their hands are visible from a distance, but this person's face may be turned away from the infant or occluded (see examples in Figure 1). Indeed, using the same pose detector on videos from in-lab play sessions, @sanchez2018postural found the opposite trend: slightly fewer hand detections than face detections from 8-16 months of age. Work that directly examines the variability in the social information in view across more vs. less structured activity contexts could further test this idea.

@@

We also used these detailed pose annotations to explore finer-grained changes in how children experience the faces and hands of their caregivers over development. We found that the faces, hands, and people in the infant view tended to become smaller and that faces tended to be farther away and in more variable horizontal positions, in keeping with prior work examining the sizes of faces in the infant view during the first year of life [@Jayaraman2015]. Overall, these data support the idea that the social information in view changes across development as infants become increasingly mobile and independent [@franchak2018; @fausey2016]. As children explore the world on their own [@xu2019towards], they may experience fewer close-up interactions with the caregivers and more bouts of play where they are exploring the objects and things in the environment around them.

More broadly, however, these analyses underscore the importance of how, when, from whom, and what data we sample; these choices become central when we attempt to draw conclusions about the regularities of experience. Indeed, while unprecedented in size, this dataset still has many limitations. These videos only represent a small portion of the everyday experience of these three children, all of whom come from relatively privileged households in western societies and thus are not representative in many ways of the global population [@karasik2018ties; @henrich2010most]. Any idiosyncrasies in how and when these particular families chose to film these videos also undoubtedly influences the variability seen here, and may contribute to the individual differences between the three children in this dataset. And without eye-tracking data, we do not know if children are attending to the social information in their visual field.

Nonetheless, we believe that these advances in datasets and methodologies represent a step in the right direction. The present paper demonstrates the feasibility of using a modern computer vision model to annotate the entirety of a very large dataset (here, >40M million frames) for the presence and size of people, hands, and faces, representing orders of magnitude more data relative to human annotations in prior work. While OpenPose did not provide annotations that were as accurate as those provided by human annotators, we found relatively consistent results with prior literature, suggesting that the sheer scale and density of the annotations provided by this method may overcome some of it's limitations.

In future work, the adaptation of deep neural networks for the infant egocentric view remains a promising avenue for collaboration between computer vision experts and developmental psychologists, and indeed has already yielded new insights about the learning mechanisms needed to build visual representations [@tsutsui2020computational; @zhuang2020unsupervised; @orhan2020self]. We propose that the use of novel algorithms with large-scale analysis of dense datasets -- collected with different fields of view, cameras, and from many different laboratories -- will lead to generalizable conclusions about the regularities of infant experience that scaffold learning.
